Automated Dynamic Slice Backsolving

### 2014-03-24 docs/paper/fsuthesis/thesis-template/paper.tex
done with paper for the day

%TODO add code examples?

@@

\section{TODO Mayhem, AEG, Exploitable Bugs}

%TODO AEG [I]?

@@

AEG required source code, where Mayhem needs access only a binary to execute. Mayhem is, however, limited to a subset of all system calls for the operating system it executes on. This is deliberate on the part of the authors, who reason that OS-wide symbolic execution introduces too much overhead. Of significance as an area of future research, Mayhem, and most other fuzzing technology explored, cannot reason about taint propagated through threading mechanisms [H].

@@

Several valuable techniques for improving fuzzers' code coverage have evolved over the years. Two, in particular, inspire this work directly: Dynamic Taint Analysis and Symbolic Execution. Dynamic Taint Analysis, or DTA, provides valuable information for guided fuzzing. Symbolic execution improves code coverage by first running a program concretely to gather constraints on a particular element, then executes the program symbolically, giving these constraints over to some kind of SMT solver [CITE] to solve for what value the element should have under certain conditions. Concolic execution, which has spawned out of the desire to use symbolic execution more wisely due to performance constraints, aims for reasonable runtimes while maintaining good code coverage. These each come with their own drawbacks, a primary one of which is the lack of automated instrumentation. Historically, any amount of symbolic execution required a programmer to directly instrument source code, leading much current research to focus on automating as much as possible.

@@

We can record taint at varying levels of granularity, with the tradeoff of finer-grained taint requiring greater system resources and slowing down analysis. Different types of taint may be more interesting to different kinds of fuzzers because different input comes from sources of varying trustedness [C]. For instance, we should probably trust raw, unsanitized user input we have read less than we trust a configuration file that only privileged users can modify. Both might be read in as input to a program, but the former is much more likely to contain malicious data like shellcode trying to exploit a vulnerability. Fuzzers have continued to drill down into finer and finer levels of granularity, from assigning everything with a generic binary value of tainted/not tainted to byte and even bit-level marking of which part of user input affected which variable. [CITE] [C] Note that this also expands taint's data type from a bit to something larger, and each tainted object requires this tag.

@@

TODO TODO TODO [C] considers concolic execution to be when we first execute concretely and THEN execute the same thing symbolically. Mixed execution is when we execute once with both concrete and symbolic input--make sure other works stand by this distinction.

TODO talk about how this is called many different things, and is a rather amorphous term.

@@

TODO why intel pin and not dynamo rio, valgrind, etc.

@@

\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [block] (driver) {driver};
    \node [cloud, left of=driver, node distance=3cm] (binary) {binary};
    \node [cloud, right of=driver, node distance=3cm] (test case) {test case};
    \node [block, below of=driver, node distance=4cm] (Pintool) {Pintool};

    %\node [block, left of=Pintool, node distance=4cm] (input) {input/};    
    %\node [block, below of=Pintool, below of=input, node distance=2cm] (output) {output/};
    %\node [block, right of=output, node distance=4cm] (explored) {explored/};
	\node [block, left of=Pintool, node distance=4cm] (results) {results};

@@

    % Draw edges
    \path [line] (driver) -- (Pintool);
    %\path [line] (Pintool) -- (input);
    %\path [line] (Pintool) -- (output);
    %\path [line] (Pintool) -- (explored);
	\path [line] (Pintool) -- (results);    
    \path [line] (Pintool) -- (decide);
    \path [line] (decide) -- node {yes} (mutate);
    \path [line] (mutate) |- (test case);
    \path [line] (decide) -- node {no}(stop);
    \path [line,dashed] (binary) -- (driver);
    \path [line,dashed] (test case) -- (driver);
    %\path [line, dashed] (explored) -- (decide);

@@

% TODO Binarys, RE?
% TODO Pin's API more details?

@@

To test Stubble's ability to perform an oracle attack, we prepared C binaries which checked user input against a secret. After compiling with the Tiny C Compiler, which performs minimal optimization and doesn't add security measures like stack cookies, this binary and a default test file are given as input to Stubble.

@@

\section{Testing}

@@

Some of the work involved in creating and testing Stubble feels spiritually similar to Netzob, a protocol reverse engineering tool which actively and passively infers protcol grammars and state machines [AQ]. Dynamic slice backsolving could be used to reverse engineer an unknown protcol by driving execution down increasingly newer and deeper paths. In order to do this, we must, by definition, more closely approximate valid input with each run, eventually mutating from random bytes to a valid protocol input.

TODO ways this technique could be used.

@@

TODO better "vuln disc." conditions like segfaults, taint reaching EIP.

>>>>

@@

\section{Mayhem, Exploitable Bugs}

@@

AEG [I], a previous work into automatically generating exploits given an exploitable condition, required source code, where Mayhem needs access only a binary to execute. Mayhem is, however, limited to a subset of all system calls for the operating system it executes on. This is deliberate on the part of the authors, who reason that OS-wide symbolic execution introduces too much overhead. Of significance as an area of future research, Mayhem, and most other fuzzing technology explored, cannot reason about taint propagated through threading mechanisms [H].

@@

Several valuable techniques for improving fuzzers' code coverage have evolved over the years. Two, in particular, inspire this work directly: Dynamic Taint Analysis and Symbolic Execution. Dynamic Taint Analysis, or DTA, provides valuable information for guided fuzzing. Symbolic execution improves code coverage by first running a program concretely to gather constraints on a particular element, then executes the program symbolically, giving these constraints over to some kind of SMT solver to solve for what value the element should have under certain conditions. Concolic execution, which has spawned out of the desire to use symbolic execution more wisely due to performance constraints, aims for reasonable runtimes while maintaining good code coverage. These each come with their own drawbacks, a primary one of which is the lack of automated instrumentation. Historically, any amount of symbolic execution required a programmer to directly instrument source code, leading much current research to focus on automating as much as possible.

@@

We can record taint at varying levels of granularity, with the tradeoff of finer-grained taint requiring greater system resources and slowing down analysis. Different types of taint may be more interesting to different kinds of fuzzers because different input comes from sources of varying trustedness [C]. For instance, we should probably trust raw, unsanitized user input we have read less than we trust a configuration file that only privileged users can modify. Both might be read in as input to a program, but the former is much more likely to contain malicious data like shellcode trying to exploit a vulnerability. Fuzzers have continued to drill down into finer and finer levels of granularity, from assigning everything with a generic binary value of tainted/not tainted to byte-level marking of user input [C]. Note that this also expands taint's data type from a "tainted-not tainted" bit to a something type, and each tainted object requires this tag.

@@

% TODO TODO TODO [C] considers concolic execution to be when we first execute concretely and THEN execute the same thing symbolically. Mixed execution is when we execute once with both concrete and symbolic input--make sure other works stand by this distinction.

% TODO talk about how this is called many different things, and is a rather amorphous term.

@@

% TODO why intel pin and not dynamo rio, valgrind, etc.

@@

\begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [block] (driver) {driver};
    \node [cloud, left of=driver, node distance=3cm] (binary) {binary};
    \node [cloud, right of=driver, node distance=3cm] (test case) {test case};
    \node [block, below of=driver, node distance=4cm] (Pintool) {Pintool};
	\node [block, left of=Pintool, node distance=4cm] (results) {results};

@@

    % Draw edges
    \path [line] (driver) -- (Pintool);
	\path [line] (Pintool) -- (results);    
    \path [line] (Pintool) -- (decide);
    \path [line] (decide) -- node {yes} (mutate);
    \path [line] (mutate) |- (test case);
    \path [line] (decide) -- node {no}(stop);
    \path [line,dashed] (binary) -- (driver);
    \path [line,dashed] (test case) -- (driver);

@@

% TODO Pin's API more details?

@@

To test Stubble's ability to perform an oracle attack, we prepared C binaries which checked user input against a secret. After compiling with the Tiny C Compiler, which performs minimal optimization and doesn't add security measures like stack cookies, this binary and a default test file are given as input to Stubble. Secrets ranged in size from 1 to 50 characters. We started with secret of one character, embedded in an arbitrary position somewhere within the several hundred byte input file. To decide how to implement dynamic slice backsolving at the assembly level, we began by statically disassembling our binaries with IDA Pro [AH]. The ability to see control flow and instructions at a block level was invaluable during testing and development. After succeeding with smaller inputs we gradually increased the size of the secret.

@@

@@

% TODO more ways DSB and stubble could be used

Dynamic Slice Backsolving would be useful when source code is not accessible, or when a user doesn't wish to instrument source code to symbolically execute a program. Other taint trackers and fuzzers could implement this technique, regardless of whether or not they require source code, to improve path exploration and help create useful mutations.

Stubble could prove useful when analysts have access to a binary, but not its source code. Running Stubble on such a binary would produce tests which will execute different paths within the binary. These tests could be used as part of a test suite, to ensure that as many paths as possible are exercised during testing.

Some of Stubble's functionality feels spiritually similar to Netzob, a protocol reverse engineering tool which actively and passively infers protcol grammars and state machines [AQ]. Dynamic slice backsolving could be used to reverse engineer an unknown protcol by driving execution down increasingly newer and deeper paths. By doing this, we must, by definition, more and more closely approximate valid input with each run, eventually mutating from random bytes to a valid protocol input. As an example, if an unknown protocol begins messages with an arbitrary string of bytes "ABC", an initial run would fail but notice that, had byte zero been "A", then a new branch could be explored. The second run would ensure that byte zero was "A", then similarly notice that had byte two been "B", yet a new branch would have been explored, and so on.

@@

There also remains research to be done regarding parallel taint tracking. We parallelized stubble at the process level, which resulted in speed improvements commensurate with the number of processes used. However, because of data dependencies, our naive implementation would not always discover and backsolve more complicated paths. We suspect solving this would require up-front analysis of data dependencies between bytes of user input before each fork off.

%Stubble would also benefit from better algorithms for discovering vulnerable conditions. Monitoring for segfaults and track taint to the instruction pointer, are good indicators, but we could also watch for memory accesses out of declared array boundaries and instrument unsafe C functions to monitor, for example, when 
%heap overflows, etc...
